---
title: "Week9-AE-RMarkdown"
output:
  pdf_document: default
  html_document: default
date: "2024-03-27"
---

```{r}
library(tidyverse)
library(readxl)
library(scales)
#install.packages('janitor')
library(janitor)
```

```{r}
nurses <- read_csv("nurses.csv") |> clean_names()

# subset to three states
nurses_subset <- nurses |>
  filter(state %in% c("California", "New York", "North Carolina"))
```

The following code chunk demonstrates how to add alternative text to a bar chart. The alternative text is added to the chunk header using the `fig-alt` chunk option. The text is written in Markdown and can be as long as needed. Note that fig-cap is not the same as `fig-alt`.

```{r}
#| label: nurses-bar
#| fig-cap: "Total employed Registered Nurses"
#| fig-alt: "The figure is a bar chart titled 'Total employed Registered
#| Nurses' that displays the numbers of registered nurses in three states
#| (California, New York, and North Carolina) over a 20 year period, with data
#| recorded in three time points (2000, 2010, and 2020). In each state, the
#| numbers of registered nurses increase over time. The following numbers are
#| all approximate. California started off with 200K registered nurses in 2000,
#| 240K in 2010, and 300K in 2020. New York had 150K in 2000, 160K in 2010, and
#| 170K in 2020. Finally North Carolina had 60K in 2000, 90K in 2010, and 100K
#| in 2020."

nurses_subset |>
  filter(year %in% c(2000, 2010, 2020)) |>
  ggplot(aes(x = state, y = total_employed_rn, fill = factor(year))) +
  geom_col(position = "dodge") +
  scale_fill_viridis_d(option = "E") +
  scale_y_continuous(labels = label_number(scale = 1/1000, suffix = "K")) +
  labs(
    x = "State", y = "Number of Registered Nurses", fill = "Year",
    title = "Total employed Registered Nurses"
  ) +
  theme(
    legend.background = element_rect(fill = "white", color = "white"),
    legend.position = c(0.85, 0.75)
    )
```

# Task 1. Add alt text to line chart
```{r}
# Your label here
#| label: nurses-salary
# Your caption here
#| fig-cap: "Annual median salary of Registered Nurses"
# Your alt text here
#| fig-alt: "The figure is a line chart titled 'Annual median salary of Registered Nurses' that displays #| the median salary of registered nurses in three states
#| (California, New York, and North Carolina) from 1998 to 2020. In each state, the
#| median salary of registered nurses increase over time. The following numbers are
#| all approximate. California nurses have the highest median salary, starting from $50K in 1998 to nearly $120K in 2020. Behind them are New York nurses, whose median salary rose from around $49K to $90K over the period. Nurses in North Carolina have the lowest median salary, going from $39K in 1998 to $67K in 2020."

nurses_subset |>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line() +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in")
    )
```

# Task 2. Direct labelling instead of legends

Create a version of the same line chart but using direct labels instead of legends. 

You can achieve this with just `geom_text()` but you can also check out https://r-graph-gallery.com/web-line-chart-with-labels-at-end-of-line.html for a fancier way of achieving this.

```{r}

library(ggrepel)

# YOUR CODE HERE
# Your label here
#| label: nurses-salary
# Your caption here
#| fig-cap: "Annual median salary of Registered Nurses"
# Your alt text here
#| fig-alt: "The figure is a line chart titled 'Annual median salary of Registered Nurses' that displays #| the median salary of registered nurses in three states
#| (California, New York, and North Carolina) from 1998 to 2020. In each state, the
#| median salary of registered nurses increase over time. The following numbers are
#| all approximate. California nurses have the highest median salary, starting from $50K in 1998 to nearly $120K in 2020. Behind them are New York nurses, whose median salary rose from around $49K to $90K over the period. Nurses in North Carolina have the lowest median salary, going from $39K in 1998 to $67K in 2020."

nurses_subset %>%
  mutate(labels = ifelse(year == 2020, state, NA))|>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(
    aes(label = labels),
    direction = "y",
    xlim = c(2022, NA),
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = .8,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20,
    show.legend = FALSE
  ) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in")
    )
```

# Task 3. Colorblind-friendly plots
Use `colorblindr` for colorblind-friendly palettes.
```{r}
# remotes::install_github("wilkelab/cowplot")
# install.packages("colorspace", repos = "http://R-Forge.R-project.org")
# remotes::install_github("clauswilke/colorblindr")
library(colorblindr)
```

Try out colorblind simulations at http://hclwizard.org/cvdemulator/ or |> your plot to `cvd_grid()` to see the plot in various color-vision-deficiency simulations.

With the line chart from Task 1, create 3 different plots: one with the default color scale, one with the `viridis` color scale, and one with the `OkabeIto` color scale from `colorblindr`. Show the `cvd_grid()` of each plot and describe the simulated effectiveness of the color scales for colorblind viewers.

```{r}
# YOUR CODE HERE for default color scale
# plot <-nurses_subset |>
#   ggplot(aes(x = year, y = annual_salary_median, color = state)) +
#   geom_line() +
#   scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
#   labs(
#     x = "Year", y = "Annual median salary", color = "State",
#     title = "Annual median salary of Registered Nurses"
#   ) +
#   coord_cartesian(clip = "off") +
#   theme(
#     plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in")
#     )
# cvd_grid(plot)

plot <- nurses_subset %>%
  mutate(labels = ifelse(year == 2020, state, NA))|>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(
    aes(label = labels),
    direction = "y",
    xlim = c(2022, NA),
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = .4,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20,
    show.legend = FALSE
  ) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in"),
    legend.position = "none"
    )
cvd_grid(plot)
```

What do you think of the default color scale effectiveness for colorblind viewers?

YOUR ANSWER HERE

Not very effective. It's hard to differentiate the lines for California and New York, especially  in the cases of deutanomaly and desaturated.

---

```{r}
# YOUR CODE HERE for viridis color scale
plot <- nurses_subset %>%
  mutate(labels = ifelse(year == 2020, state, NA))|>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(
    aes(label = labels),
    direction = "y",
    xlim = c(2022, NA),
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = .4,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20,
    show.legend = FALSE
  ) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in"),
    legend.position = "none"
    ) +
  scale_color_viridis_d()
cvd_grid(plot)
```

What do you think of the viridis color scale effectiveness for colorblind viewers?

YOUR ANSWER HERE

Still quite hard to differentiate between California's line and New York's line; North Carolina is hard to see overall. 

---

```{r}
# YOUR CODE HERE for OkabeIto color scale
plot <- nurses_subset %>%
  mutate(labels = ifelse(year == 2020, state, NA))|>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(
    aes(label = labels),
    direction = "y",
    xlim = c(2022, NA),
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = .4,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20,
    show.legend = FALSE
  ) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in"),
    legend.position = "none"
    ) +
  scale_color_OkabeIto()
cvd_grid(plot)
```

What do you think of the OkabeIto color scale effectiveness for colorblind viewers? 

YOUR ANSWER HERE

Easier to differentiate in all cases except perhaps desaturated.

