---
title: "COMP4010 - Week 2"
output:
  pdf_document: default
  html_document: default
date: "2024-02-28"
---

# Week 2

# Application Exercises

Include `tidyverse`:

```{r}
#install.packages("tidyverse")
library(tidyverse)
theme_set(theme_minimal())
```

Read the data:

```{r}
df <- read_csv("homesales.csv")
```

Average home size by decade:

```{r}
# create decade variable
df <- df |>
  mutate(
    decade_built = (year_built %/% 10) * 10,
    decade_built_cat = case_when(
      decade_built <= 1940 ~ "1940 or before",
      decade_built >= 1990 ~ "1990 or after",
      .default = as.character(decade_built)
    )
  )

# calculate mean area by decade
mean_area_decade <- df |>
  group_by(decade_built_cat) |>
  summarize(mean_area = mean(area))
mean_area_decade
```

Visualizing the data as a bar chart:

```{r}
ggplot(
  data = mean_area_decade,
  mapping = aes(x = mean_area, y = decade_built_cat)
) +
  geom_col() +
  labs(
    x = "Mean area (square feet)", y = "Decade built",
    title = "Mean area of houses, by decade built"
  )
```

Visualizing the data as a dot plot:

```{r}
ggplot(
  data = mean_area_decade,
  mapping = aes(x = mean_area, y = decade_built_cat)
) +
  geom_point(size = 4) +
  labs(
    x = "Mean area (square feet)", y = "Decade built",
    title = "Mean area of houses, by decade built"
  )
```

## TASK 1. Visualizing the data as a lollipop chart

```{r}
# YOUR CODE HERE
ggplot(
  data = mean_area_decade,
  mapping = aes(x = mean_area, y = decade_built_cat)
) +
  geom_segment(aes(x = 0, xend = mean_area, y = decade_built_cat, yend = decade_built_cat)) +
  geom_point() +
  labs(
    x = "Mean area (square feet)", y = "Decade built",
    title = "Mean area of houses, by decade built"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))



```

## TASK 2. Visualizing the distribution of the number of bedrooms

Collapse the variable `beds` into a smaller number of categories and drop rows with missing values for this variable:

```{r}
df_bed <- df |>
  mutate(beds = factor(beds) |>
    fct_collapse(
      "5+" = c("5", "6", "7", "9")
    )) |>
  drop_na(beds)
```

```{r}
# YOUR CODE HERE
ggplot(data = df_bed, aes(x = beds, fill = beds)) + geom_bar() +
labs(x = "Number of bedrooms", y = "Count",
title = "Distribution of the number of bedrooms") + 
theme(plot.title = element_text(hjust = 0.5))
```

## TASK 3. Visualizing the distribution of the number of bedrooms by the decade in which the property was built

Stacked bar chart (number of bedrooms by the decade built):

```{r}
# YOUR CODE HERE

ggplot(data = df_bed, aes(x = decade_built_cat, fill = beds)) + geom_bar() +
labs(x = "Decade built", y = "Properties sold",
title = "Distribution of the number of bedrooms by decade") + 
theme(plot.title = element_text(hjust = 0.5))


```

Dodged bar chart (number of bedrooms by the decade built):

```{r}
# YOUR CODE HERE

ggplot(data = df_bed, aes(x = decade_built_cat, fill = beds)) + geom_bar(position = "dodge") +
labs(x = "Decade built", y = "Properties sold",
title = "Distribution of the number of bedrooms by decade") + 
theme(plot.title = element_text(hjust = 0.5))


```

Relative frequency bar chart (number of bedrooms by the decade built):

```{r}
# YOUR CODE HERE
ggplot(data = df_bed, aes(x = decade_built_cat, fill = beds)) + geom_bar(position = "fill") +
labs(x = "Decade built", y = "Properties sold",
title = "Distribution of the number of bedrooms by decade") + 
theme(plot.title = element_text(hjust = 0.5))



```

## Task 4. Visualizing the distribution of property size by decades

Getting mean of area of each decade category:

```{r}
mean_area_decade <- df |>
  group_by(decade_built_cat) |>
  summarize(mean_area = mean(area))
```

Bar chart (mean area by decade built):

```{r}
# YOUR CODE HERE

ggplot(data = mean_area_decade, aes(x = decade_built_cat, y = mean_area)) + geom_bar(stat = "identity") +
labs(x = "Decade built", y = "Mean area",
title = "Mean area by decade") + 
theme(plot.title = element_text(hjust = 0.5))


```

Box plot (area by decade built):

```{r}
# YOUR CODE HERE

ggplot(data = df, aes(x = decade_built_cat, y = area)) + geom_boxplot() +
labs(x = "Decade built", y = "Area",
     title = "Distribution of area by decade") + 
theme(plot.title = element_text(hjust = 0.5))



```

Violin plot (area by decade built):

```{r}
# YOUR CODE HERE

ggplot(data = df, aes(x = decade_built_cat, y = area)) + geom_violin() +
labs(x = "Decade built", y = "Area",
     title = "Distribution of area by decade") + 
theme(plot.title = element_text(hjust = 0.5))


```

Strip chart (area by decade built):

```{r}
set.seed(4010)

# YOUR CODE HERE

ggplot(data = df, aes(x = decade_built_cat, y = area)) + geom_jitter(alpha = 0.3) +
labs(x = "Decade built", y = "Area",
     title = "Distribution of area by decade") + 
theme(plot.title = element_text(hjust = 0.5))


```
